## Isolation by distance

We assessed isolation by distance by comparing genetic distance to geographic distance. Specifically, we took the traditional approach of comparing a geographic dissimilarity matrix (based on latitude and longitude) to a genetic dissimilarity matrix. We calculated the genetic dissimilarity matrix with the `dist.genpop` function int the adegenet package. We use the [Cavalli-Sforza](https://doi.org/10.2307/2406616) distance metric, or `method = 2` argument for the `dist.genpop` function.

Note that for this analysis, we treated each *sampling site* as a distinct location. There would not be enough power to do a distance matrix among 3-5 cities. Code for generating stats and figures from the Mantel test can be found in the source code below.

```{r ibd_1, message=FALSE}
source("R/isolation_by_distance.R")
```

```{r ibd_2, eval = FALSE, message=FALSE}
extract_ibd_stats_and_plots()
```

Below are the results of the mantel test. Note that there is a p-value correction for testing multiple cities (species are treated as independent, however).

```{r ibd_3, echo = FALSE, message=FALSE}
table_ <- readr::read_csv("output/IBD/isolation-by-distance-mantel-test.csv") %>% dplyr::select(-c(Hypothesis, Reps))
knitr::kable(
  table_,
  format = "simple",
  caption = "Statistics from running 9999 permutations ('Reps') via mantel test, limited to genomic versus distance comparisons. Hypothesis for all tests is 'greater'."
)
```

We also repeated this within city. Note that there is a p-value correction for testing multiple environmental variables and cities (species are treated as independent, however).

\tiny

```{r ibd_4, echo = FALSE, message=FALSE}
table_ <- readr::read_csv("output/IBD/isolation-by-distance-within-city-mantel-test.csv") %>% dplyr::select(-c(Hypothesis, Reps))
knitr::kable(
  table_,
  format = "simple",
  caption = "Statistics from running 9999 permutations ('Reps') via mantel test, limited to within city for genomic versus distance comparisons. Hypothesis for all tests is 'greater'."
)
```

\normalsize