## Isolation by environment

### Environmental data

Environmental variables include the monthly averages in the middle of the day for:

- air temperature at 5cm above ground
- air temperature at 1.2m above ground
- soil temperature at 2.5cm below ground
- RH (relative humidity) at 5cm above ground
- RH at 1.2m above ground

Variables were extracted from historic datasets and modeled using a microclimate model. More information can be found on the [NicheMapR website](https://mrke.github.io/) (how the model works, what variables can be manipulated and what you can model, vignettes for running models in R). 

This method was chosen because it takes data from global datasets (you can use both historic and current or pick specific years) but then accounts for site-specific variables (we can change the % shade, the slope or aspect of the landscape, and it considers elevation, average cloud cover, etc.). [Here’s the list](https://mrke.github.io/models/MicroClimate-Models) of all the different models/datasets we’re able to can pull from. It's meant for mechanistic niche modeling.

Variables in the file `site_data_DUC_environvars.csv` are all for the monthly averages at noon (12pm - hottest part of the day!) and are extreme. In other words, they are maximums. 

Note that [this Stack Overflow post](https://stackoverflow.com/questions/69639782/installing-gfortran-on-macbook-with-apple-m1-chip-for-use-in-r) is helpful with installing `NicheMapR`.

```{r}
# devtools::install_github('mrke/NicheMapR')
# library(NicheMapR)
# 
# test_site_coords <- c(sites[1,]$lat, sites[1,]$long)
# test_distance_to_city_center_km <- sites[1,]$distance_to_city_center_km
# micros_ <- micro_usa(loc = test_site_coords)
# 
# loc <- c(-89.40, 43.07)
# micro <- micro_global(loc = loc)
```

### IBE analysis

We assessed isolation by environment by comparing genetic distance to environmental distance, or the difference among sites. Genetic distance was generated the same way as isolation by distance (IBD) above. Code for generating stats from the Mantel test can be found in the source code below.

```{r ibe_1, message=FALSE}
source("R/isolation_by_environment.R")
```

The following functions are used to generate statistics for all environmental variables. Default of the function runs for "nlcd_urban_pct" which is the percent urban cover of a site within city. These are the four environmental variables mentioned in the main manuscript, although more environmental variables are present in the raw data.

```{r ibe_2, eval = FALSE, message=FALSE}
extract_ibe_stats_and_plots()
extract_ibe_stats_and_plots(env_var_to_use = "distance_to_city_center_km")
extract_ibe_stats_and_plots(env_var_to_use = "soiltemp_2.5cm_Jul_12pm")
extract_ibe_stats_and_plots(env_var_to_use = "soiltemp_2.5cm_Apr_12pm")
```

Below are the results of the mantel test. Note that there is a p-value correction for testing multiple environmental variables (species are treated as independent, however).

\tiny

```{r ibe_3, echo = FALSE, message=FALSE}
table_ <- readr::read_csv("output/IBE/isolation-by-env-all-mantel-test.csv") %>% dplyr::select(-c(Hypothesis, Reps))
knitr::kable(
  table_,
  format = "simple",
  caption = "Statistics from running 9999 permutations ('Reps') via mantel test, for genomic versus environmental comparisons. Hypothesis for all tests is 'greater'."
)
```

\normalsize

We also repeated the mantel tests within city. Note that there is a p-value correction for testing multiple environmental variables and cities (species are treated as independent, however).

\tiny

```{r ibe_5, echo = FALSE, message=FALSE}
table_ <- readr::read_csv("output/IBE/isolation-by-env-within-city-mantel-test.csv") %>% dplyr::select(-c(Hypothesis, Reps))
knitr::kable(
  table_,
  format = "simple",
  caption = "Statistics from running 9999 permutations ('Reps') via mantel test, limited to within city for genomic environmental comparisons. Hypothesis for all tests is 'greater'."
)
```

\normalsize

The following function makes the main text figure.

```{r eval = FALSE}
ibe_mega_plot()
```
